---
title: "homework2"
author: "Claire Whittington"
date: "2023-10-06"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)      # Our usual data manipulation tools
library(readxl)         # For reading MS Excel files
library(googlesheets4)  # For reading Google Sheets
```
## (1) Exercise 1 – 5 pts. In exercise 1:
• (1 pts) Import the correct sheet from Example_5, without modifying it in any way.
```{r}
# grab sheet 2 and put it into my.trees
my.trees <- read_excel("~/Downloads/Example_5.xls", sheet = 2)
my.trees
```

• ( ½ pt) How many blank columns were imported?

- Two blank columns


• ( ½ pt) Show the structure of the structure of the data using str(yourdataframe’sname)
```{r}
str(my.trees)
```

• (3 pts) Modify the data frame so that it only has 3 columns of 31 observations. Rename
the columns to have the appropriate names of “Girth”, “Height,” and “Volume” if your
step above did not automatically rename the columns.
```{r}
# remove rows 1,2,3 and cols 4 and 5
cleaned_data <- my.trees[-c(0:4),-c(4,5)] 

# rename cols
colnames(cleaned_data)[1] <- "Girth"
colnames(cleaned_data)[2] <- "Height"
colnames(cleaned_data)[3] <- "Volume"

# show structure summary
str(cleaned_data)

# show data
head(cleaned_data)
```
• Your output should have the answer to question in the 2nd bulletin point, the structure of
the data frame, and the final data frame. Include all code used to clean up the data frame.

## (2) Exercise 2 – 4 pts. In exercise 2,
• (1 pts) Import the correct sheet from Example_5, without modifying it in any way.
```{r}
my.cars <- read_excel("~/Downloads/Example_3.xls", sheet = 2)
head(my.cars)
```

• (2 pts) Use View(yourdataframe’sname). Write 2 – 3 sentences contrasting what you saw
in the Excel file and what the data frame looks like to R.
```{r}
head(my.cars)
```
In the excel file, I just see 12 columns all properly named with their 
corresponding data. In R, there are 82 columns and like 1600 rows with NA values

• (1 pt) Modify the data frame so that it only has the 12 columns x 33 rows of data.
```{r}
cleaned_data <- my.cars[1:33,1:12]
cleaned_data
```
(3) Turned in by the due date/time – 1 point