---
title: "homework7"
author: "Claire Whittington"
date: "2023-11-06"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)   # loading ggplot2 and dplyr
library(patchwork)   # arranging multiple graphs into 1 figure
library(viridis)     # The viridis color schemes
library(latex2exp)   # For plotting math notation
library(plotly)     # for interactive hover-text
library(faraway)
```

Please complete:

## (1) Exercise 1 – 6 pts

The infmort data set from the package faraway gives the infant mortality rate for a variety of countries. The information is relatively out of date (from 1970s?), but will be fun to graph. 

Visualize the data using by creating scatter plots of mortality vs income while faceting using region and setting color by oil export status. 

Utilize a log10 transformation for both mortality and income axes. 

This can be done either by doing the transformation inside the aes() command or by utilizing the scale_x_log10() or scale_y_log10() layers. 

The critical difference is if the scales are on the original vs log transformed scale. Experiment with both and see which you prefer.
```{r}
head(infmort)
```

a ) The rownames() of the table gives the country names and you should create a new column that contains the country names. *rownames
```{r}
infmort <- infmort %>% 
            mutate ( country = rownames(infmort) )
head(infmort)
```

b) Create scatter plots with the log10() transformation inside the aes() command.
```{r}
ggplot( infmort, aes( x = log10(mortality), y = log10(income), color = oil)) +
  geom_point(size = 3) +
  facet_wrap(~region)
```

c) Create the scatter plots using the scale_x_log10() and scale_y_log10(). Set the major and minor breaks to be useful and aesthetically pleasing. Comment on which version you find easier to read.

COMMENT: I don't believe any major or minor tick mark adjustments are needed.
Europe's highest income is over 3000, and Asia's mortality is over 1000.
I don't believe we need to scale up or down our tick marks to effectively present
this data, especially since we are using log10 transform. However, I like using 
the scale command after we set the un-adjusted x and y axises, because it feels
much more modular
```{r}

ggplot( infmort, aes( x = mortality, y = income, color = oil)) +
  geom_point(size = 3) +
  facet_wrap(~region) +
  scale_x_log10() + 
  scale_y_log10() 
  #scale_x_continuous( breaks=seq(0, 1000, by=100), minor_breaks=0:1000)
```


d) The package ggrepel contains functions geom_text_repel() and geom_label_repel() that mimic the basic geom_text() and geom_label() functions in ggplot2, but work to make sure the labels don’t overlap. Select 10-15 countries to label and do so using the geom_text_repel() function.
```{r}
library(ggrepel)
selected_countries <- sample(infmort$country, 15)
selected_data <- infmort[infmort$country %in% selected_countries, ]
head(selected_data)

ggplot( infmort, aes( x = mortality, y = income, color = oil)) +
  geom_point(size = 3) +
  geom_text_repel(data = selected_data, aes(label = country),
                  size = 3.2, box.padding = 0.5) +
  facet_wrap(~region) +
  scale_x_log10() + 
  scale_y_log10() 

```

## (2) Exercise 2 – 7 pts
Using the datasets::trees data, complete the following:

a) Create a regression model for  y= Volume as a function of  x = Height.
```{r}
model <- lm(Volume ~ Height, trees)
```


b) Using the summary command, get the y-intercept and slope of the regression line.
```{r}
summary(model)

intercept <- coef(model)[1]
slope <- coef(model)[2]
```

c) Using ggplot2, create a scatter plot of Volume vs Height.

```{r}
P <- ggplot( trees, aes(x = Height, y=Volume) ) + geom_point()
P
```

d) Create a nice white filled rectangle to add text information to using by adding the following annotation layer.
```{r}
NP <- P + annotate('rect', xmin=65, xmax=75, ymin=60, ymax=74,
         fill='white', color='black',)
NP 
```

e) Add some annotation text to write the equation of the line  <equation in book>  in the text area.
```{r}
lm_line = "Yi = -87.12 + 1.54 * xi"
NP <- NP + annotate("text", x = 70, y = 67, label = lm_line, col = "black", cex = 6)
NP
```

f) Add annotation to add R2 = 0.358
```{r}
NP <- NP + annotate("text", x = 70, y = 62, label = "R2 = 0.358", col = "black", cex = 3.7)
NP
```

g) Add the regression line in red. The most convenient layer function to uses is geom_abline(). It appears that the annotate doesn’t work with geom_abline() so you’ll have to call it directly
```{r}
LMP <- NP +
      geom_abline(intercept = intercept, slope = slope, color = "red")
LMP
```


(3) Exercise 3 – Skip: Come back to it later if you are interested.

(7) Turned in by the due date/time – 2 points. More than a week late = 0 points for the
assignment